### 2024-03-29 doc/sections/mav-cut-elimination.tex
Consistent use of structure

Let $\Chu(\ClosedLowerSet{\NMAV}, \ClosedDay{I})$ be the MAV-algebra
constructed (\Cref{thm:algebra-from-frame}) from the normal
proof MAV-frame (\Cref{prop:nmav-frame}), where elements
are positive/negative pairs of $+$-closed lower sets of formulas. We
define the valuation of atoms as
$V(\va) = \ChuEmbed(\ClosedLowerEmbed(\vDual\va))$. By
\Cref{thm:soundness}, we have an interpretation of MAV formulas
$\sem{\vP}$ such that if $\vP\vInferFrom*\vUnit$, then
$I \sqsubseteq \sem{\vP}$. We now prove our main proposition about this
interpretation in the MAV-algebra derived from the MAV-frame of normal
proofs \NMAV{} that will allow us to derive the admissibility of all
the non-normal proof rules of MAV.

>>>>

Let $\Chu(\ClosedLowerSet{\NMAV}, \ClosedDay{I})$ be the MAV-algebra
constructed (\Cref{thm:algebra-from-frame}) from the normal
proof MAV-frame (\Cref{prop:nmav-frame}), where elements
are positive/negative pairs of $+$-closed lower sets of structures. We
define the valuation of atoms as
$V(\va) = \ChuEmbed(\ClosedLowerEmbed(\vDual\va))$. By
\Cref{thm:soundness}, we have an interpretation of MAV structures
$\sem{\vP}$ such that if $\vP\vInferFrom*\vUnit$, then
$I \sqsubseteq \sem{\vP}$. We now prove our main proposition about this
interpretation in the MAV-algebra derived from the MAV-frame of normal
proofs \NMAV{} that will allow us to derive the admissibility of all
the non-normal proof rules of MAV.

### 2024-03-29 doc/sections/mav-mechanisation.tex
Consistent use of structure

As well as giving confidence in our results, a key benefit of
formalising the development in a proof assistant like Agda is that the
proof normalisation process (\Cref{thm:cut-elim}) is
executable. For example, we can normalise the one step SMAV proof
(using \cref{rule:Axiom}) of $\vP \vParr \vDual{\vP}$ where
$\vP = (\vUnit \vPlus \vUnit) \vSeq (\vUnit \vWith \vUnit)$ to a FIXME
step MAV proof, given in the Appendix.

>>>>

As well as giving confidence in our results, a key benefit of
formalising the development in a proof assistant like Agda is that the
proof normalisation process (\Cref{thm:cut-elim}) is
executable. For example, we can normalise the one step SMAV proof
using \cref{rule:Axiom} of $\vP \vParr \vDual{\vP}$ where
$\vP = (\vUnit \vPlus \vUnit) \vSeq (\vUnit \vWith \vUnit)$ to a FIXME
step MAV proof, given in the Appendix.

### 2024-03-29 doc/sections/mav-syntax.tex
Adjust positive and negative atoms to match use in semantics
Add index and restructure inference rule table
Consistent use of structure

      l@{\;=\;}l @{\hspace{1cm}}
      l@{\;=\;}l @{\hspace{1cm}}
      l@{\;=\;}l @{\hspace{1cm}}
      l@{\;=\;}l}
    \vDual{\vPos\va}     & \vNeg\va
    \vDual{\vUnit}       & \vUnit
    \vDual{\vP\vTens\vQ} & \vDual\vP \vParr \vDual\vQ
    \vDual{\vP\vParr\vQ} & \vDual\vP \vTens \vDual\vQ

@@

    \vDual{\vNeg\va}     & \vPos\va
    \vDual{\vP\vSeq \vQ} & \vDual\vP \vSeq  \vDual\vQ
    \vDual{\vP\vWith\vQ} & \vDual\vP \vPlus \vDual\vQ
    \vDual{\vP\vPlus\vQ} & \vDual\vP \vWith \vDual\vQ

@@

\emph{Inference}, written $\vInferFrom$, is the smallest relation defined by the following axioms:
% NOTE: These shenanigans are for cross-text table alignment.
\newlength{\vInferFromDefnWidthOfLBL}%
\settowidth{\vInferFromDefnWidthOfLBL}{$\RuleName{AtomAxiom}$}%
\newlength{\vInferFromDefnWidthOfLHS}%
\settowidth{\vInferFromDefnWidthOfLHS}{$(\vP\vSeq\vQ)\vParr(\vR\vSeq\vS)$}%
\newlength{\vInferFromDefnWidthOfRHS}%
\settowidth{\vInferFromDefnWidthOfRHS}{$(\vP\vParr\vR)\vWith(\vQ\vParr\vR)$}%
\begin{displaymath}
  \begin{array}{lcl@{\hspace{0.5cm}}|@{\hspace{0.5cm}}l}
    \vP\vParr\vDual\vP
     & \vInferFrom
     & \vUnit
     & \makebox[\vInferFromDefnWidthOfLBL][l]{$\RuleLabel{Axiom}$}
    \\
    \vUnit
     & \vInferFrom
     & \vP\vTens\vDual\vP
     & \RuleLabel{Cut}
    \\
    (\vP\vTens\vQ)\vParr\vR
     & \vInferFrom
     & \vP\vTens(\vQ\vParr\vR)
     & \RuleLabel{Switch}
    \\
    \vUnit\vWith\vUnit
     & \vInferFrom
     & \vUnit
     & \RuleLabel{Tidy}
    \\
    \vUnit
     & \vInferFrom
     & \vUnit \vPlus \vUnit
     & \RuleLabel{CoTidy}
    \\
    (\vP\vSeq\vQ)\vParr(\vR\vSeq\vS)
     & \vInferFrom
     & (\vP\vParr\vR)\vSeq(\vQ\vParr\vS)
     & \RuleLabel{Sequence}
    \\
    (\vP\vTens\vR)\vSeq(\vQ\vTens\vS)
     & \vInferFrom
     & (\vP\vSeq\vQ)\vTens(\vR\vSeq\vS)
     & \RuleLabel{CoSequence}
    \\
    \vP\vPlus\vQ
     & \vInferFrom
     & \vP
     & \RuleLabel{Left}
    \\
    \vP
     & \vInferFrom
     & \vP\vWith\vQ
     & \RuleLabel{CoLeft}
    \\
    \vP\vPlus\vQ
     & \vInferFrom
     & \vQ
     & \RuleLabel{Right}
    \\
    \vQ
     & \vInferFrom
     & \vP\vWith\vQ
     & \RuleLabel{CoRight}
    \\
    (\vP\vWith\vQ)\vParr\vR
     & \vInferFrom
     & (\vP\vParr\vR)\vWith(\vQ\vParr\vR)
     & \RuleLabel{External}
    \\
    (\vP \vTens \vR) \vPlus (\vQ \vTens \vR)
     & \vInferFrom
     & (\vP \vPlus \vQ) \vTens \vR
     & \RuleLabel{CoExternal}
    \\
    (\vP\vSeq\vQ)\vWith(\vR\vSeq\vS)
     & \vInferFrom
     & (\vP\vWith\vR)\vSeq(\vQ\vWith\vS)
     & \RuleLabel{Medial}
    \\
    (\vP \vPlus \vR) \vSeq (\vQ \vPlus \vS)
     & \vInferFrom
     & (\vP \vSeq \vQ) \vPlus (\vR \vSeq \vS)
     & \RuleLabel{CoMedial}
    \\
    \vC\vPlug\vP\vInferFrom\vC\vPlug\vQ
     & \text{if}
     & \vP\vInferFrom\vQ
     & \RuleLabel{Mono}
    % \\
    % \vP\vInferFrom\vQ
    %  & \text{if}
    %  & \vP\vEquiv\vQ
    %  & \RuleLabel{Equiv}
  \end{array}
\end{displaymath}
(If $\vP\vInferFrom\vQ$, we say that $\vP$ can be inferred from $\vQ$, \ie the arrow points from conclusion to premise.)

Note that, except for \cref{rule:Switch}, the rules come in pairs that are dual, in the sense that if a rule is of the form $\vP\vInferFrom\vQ$, then its partner is of the form $\vDual\vP\vInferFrom\vDual\vQ$. The \cref{rule:Switch} rule is, up to commutativity, already its own dual. The \cref{rule:Cut}, \cref{rule:CoLeft}, and \cref{rule:CoRight} rules introduce of new formulas going left-to-right. Normal proofs, which we define below, avoid these \emph{synthetic} rules.

@@

\begin{remark}
  In BV, the structural connectives are usually presented as lists, distinguished only by their brackets: $\vP\vTens\vQ$ is written as $\vls(\vP;\vQ)$; $\vP\vParr\vQ$ is written as $\vls[\vP;\vQ]$; and $\vP\vSeq\vQ$ is written as $\vls<\vP;\vQ>$.
  % The equations defining duality are usually included in the decidable equality on structures.
  Inferences, derivations, and proofs are presented vertically, as (\ref{notation:vlin}), (\ref{notation:vlde}), and (\ref{notation:vlpr}), respectively.
  \begin{center}
    $\inlineequation[notation:vlin]{%
        \vlderivation{\vlin{}{}{\vP}{\vlhy{\vQ}}}}$
    \qquad
    $\inlineequation[notation:vlde]{%
        \vlderivation{\vlde{}{}{\vP}{\vlhy{\vQ}}}}$
    \qquad
    $\inlineequation[notation:vlpr]{%
        \vlderivation{\vlpr{}{}{\vP}}}$
  \end{center}
  The relation between the deductive system for BV and rewrite systems is well-known, \eg by Kahramanogullari~\cite{Kahramanogullari08:maude}, who implements proof search for several deep inference systems in Maude~\cite{ClavelDELMMQ02:maude}.
\end{remark}

\begin{definition}\label{defn:normal}
  A derivation is \emph{normal} when it does not use \cref{rule:Cut} or any of the (\textsc{Co-X}) rules and its uses of \cref{rule:Axiom} are restricted to \cref{rule:AtomAxiom}, as defined by the following axiom:
  \begin{displaymath}
    \begin{array}{lcl@{\hspace{0.5cm}}|@{\hspace{0.5cm}}l}
      \makebox[\vInferFromDefnWidthOfLHS][l]{$\vNeg\va\vParr\vPos\va$}
       & \vInferFrom
       & \makebox[\vInferFromDefnWidthOfRHS][l]{$\vUnit$}
       & \RuleLabel{AtomAxiom}
    \end{array}
  \end{displaymath}
\end{definition}
Normal derivations avoid the use of rules that introduce new formulas into proofs, and so can be termed \emph{analytic} in contrast to the need for the synthetic rules to synthesise new formulas.

Our main result, Theorem \ref{thm:cut-elim}, is that every formula that is provable in full MAV also has a normal proof. Therefore, the system with only analytic rules is complete for MAV provability. Horne \cite{Horne15:mav} proves this result via a syntactic proof involving rewriting and termination measures. In the following sections, we construct a semantic proof that normal proofs are complete for MAV.

>>>>

      lcl @{\hspace{1cm}}
      lcl @{\hspace{1cm}}
      lcl @{\hspace{1cm}}
      lcl}
    \multicolumn{3}{l}{}
    % \vDual{\vPos\va}
    %  & =
    %  & \vNeg\va
     & \vDual{\vUnit}
     & =
     & \vUnit
     & \vDual{\vP\vTens\vQ}
     & =
     & \vDual\vP \vParr \vDual\vQ
     & \vDual{\vP\vParr\vQ}
     & \vDual\vP \vTens \vDual\vQ

@@

    \vDual{\vNeg\va}
     & =
     & \vPos\va
     & \vDual{\vP\vSeq \vQ}
     & =
     & \vDual\vP \vSeq  \vDual\vQ
     & \vDual{\vP\vWith\vQ}
     & =
     & \vDual\vP \vPlus \vDual\vQ
     & \vDual{\vP\vPlus\vQ}
     & \vDual\vP \vWith \vDual\vQ

@@

\emph{Inference}, written $\vInferFrom$, is the smallest relation defined by the following axioms:
% NOTE: These shenanigans are for cross-text table alignment.
\newlength{\vInferFromDefnWidthOfLBL}%
\settowidth{\vInferFromDefnWidthOfLBL}{$\RuleName{AtomAxiom}$}%
\newlength{\vInferFromDefnWidthOfLHS}%
\settowidth{\vInferFromDefnWidthOfLHS}{$(\vP\vSeq\vQ)\vParr(\vR\vSeq\vS)$}%
\newlength{\vInferFromDefnWidthOfRHS}%
\settowidth{\vInferFromDefnWidthOfRHS}{$(\vP\vParr\vR)\vWith(\vQ\vParr\vR)$}%
\begin{displaymath}
  \begin{array}{lcl@{\hspace{0.5cm}}|@{\hspace{0.5cm}}l}
    \vP\vParr\vDual\vP
     & \vInferFrom
     & \vUnit
     & \makebox[\vInferFromDefnWidthOfLBL][l]{$\RuleLabel{Axiom}$}
    \\
    (\vP\vTens\vQ)\vParr\vR
     & \vInferFrom
     & \vP\vTens(\vQ\vParr\vR)
     & \RuleLabel{Switch}
    \\
    \vUnit\vWith\vUnit
     & \vInferFrom
     & \vUnit
     & \RuleLabel{Tidy}
    \\
    (\vP\vSeq\vQ)\vParr(\vR\vSeq\vS)
     & \vInferFrom
     & (\vP\vParr\vR)\vSeq(\vQ\vParr\vS)
     & \RuleLabel{Sequence}
    \\
    \vP\vPlus\vQ
     & \vInferFrom
     & \vP
     & \RuleLabel{Left}
    \\
    \vP\vPlus\vQ
     & \vInferFrom
     & \vQ
     & \RuleLabel{Right}
    \\
    (\vP\vWith\vQ)\vParr\vR
     & \vInferFrom
     & (\vP\vParr\vR)\vWith(\vQ\vParr\vR)
     & \RuleLabel{External}
    \\
    (\vP\vSeq\vQ)\vWith(\vR\vSeq\vS)
     & \vInferFrom
     & (\vP\vWith\vR)\vSeq(\vQ\vWith\vS)
     & \RuleLabel{Medial}
    \\
     &
     &
     &
    \\
    \vUnit
     & \vInferFrom
     & \vP\vTens\vDual\vP
     & \RuleLabel{Cut}
    \\
    \vUnit
     & \vInferFrom
     & \vUnit \vPlus \vUnit
     & \RuleLabel{CoTidy}
    \\
    (\vP\vTens\vR)\vSeq(\vQ\vTens\vS)
     & \vInferFrom
     & (\vP\vSeq\vQ)\vTens(\vR\vSeq\vS)
     & \RuleLabel{CoSequence}
    \\
    \vP
     & \vInferFrom
     & \vP\vWith\vQ
     & \RuleLabel{CoLeft}
    \\
    \vQ
     & \vInferFrom
     & \vP\vWith\vQ
     & \RuleLabel{CoRight}
    \\
    (\vP \vTens \vR) \vPlus (\vQ \vTens \vR)
     & \vInferFrom
     & (\vP \vPlus \vQ) \vTens \vR
     & \RuleLabel{CoExternal}
    \\
    (\vP \vPlus \vR) \vSeq (\vQ \vPlus \vS)
     & \vInferFrom
     & (\vP \vSeq \vQ) \vPlus (\vR \vSeq \vS)
     & \RuleLabel{CoMedial}
    \\
     &
     &
     &
    \\
    \vC\vPlug\vP\vInferFrom\vC\vPlug\vQ
     & \text{if}
     & \vP\vInferFrom\vQ
     & \RuleLabel{Mono}
    % \\
    % \vP\vInferFrom\vQ
    %  & \text{if}
    %  & \vP\vEquiv\vQ
    %  & \RuleLabel{Equiv}
  \end{array}
\end{displaymath}
If $\vP\vInferFrom\vQ$, we say that $\vP$ can be inferred from $\vQ$, \ie the arrow points from conclusion to premise.

The inference rules come in dual pairs. For every rule $\vP\vInferFrom\vQ$, there is a dual rule $\vDual\vP\vInferFrom\vDual\vQ$.
The exception is \cref{rule:Switch}, which is self-dual, up to commutativity.
The \cref{rule:Cut}, \cref{rule:CoLeft}, and \cref{rule:CoRight} rules introduce of new structures going left-to-right.
Normal proofs, which we define below, avoid these \emph{synthetic} rules.

@@

\begin{remark}
  In BV, the structural connectives are usually presented as lists, distinguished only by their brackets: $\vP\vTens\vQ$ is written as $\vls(\vP;\vQ)$; $\vP\vParr\vQ$ is written as $\vls[\vP;\vQ]$; and $\vP\vSeq\vQ$ is written as $\vls<\vP;\vQ>$.
  % The equations defining duality are usually included in the decidable equality on structures.
  Inferences, derivations, and proofs are presented vertically, as (\ref{notation:vlin}), (\ref{notation:vlde}), and (\ref{notation:vlpr}), respectively.
  \begin{center}
    $\inlineequation[notation:vlin]{%
        \vlderivation{\vlin{}{}{\vP}{\vlhy{\vQ}}}}$
    \qquad
    $\inlineequation[notation:vlde]{%
        \vlderivation{\vlde{}{}{\vP}{\vlhy{\vQ}}}}$
    \qquad
    $\inlineequation[notation:vlpr]{%
        \vlderivation{\vlpr{}{}{\vP}}}$
  \end{center}
  The relation between the deductive system for BV and rewrite systems is well-known, \eg by Kahramanogullari~\cite{Kahramanogullari08:maude}, who implements proof search for several deep inference systems in Maude~\cite{ClavelDELMMQ02:maude}.
  Inferences rules are usually named with the combination of a letter and an up or down arrow, \eg \cref{rule:Axiom} and \cref{rule:Cut} are $\iD$ and $\iU$, respectively.
\end{remark}

\begin{definition}\label{defn:normal}
  A derivation is \emph{normal} when it does not use \cref{rule:Cut} or any of the (\textsc{Co-X}) rules and its uses of \cref{rule:Axiom} are restricted to \cref{rule:AtomAxiom}, as defined by the following axiom:
  \begin{displaymath}
    \begin{array}{lcl@{\hspace{0.5cm}}|@{\hspace{0.5cm}}l}
      \makebox[\vInferFromDefnWidthOfLHS][l]{$\vNeg\va\vParr\vPos\va$}
       & \vInferFrom
       & \makebox[\vInferFromDefnWidthOfRHS][l]{$\vUnit$}
       & \RuleLabel{AtomAxiom}
    \end{array}
  \end{displaymath}
\end{definition}
Normal derivations avoid the use of rules that introduce new structures into proofs, and so can be termed \emph{analytic} in contrast to the need for the synthetic rules to synthesise new structures.

Our main result, \Cref{thm:cut-elim}, is that every structure that is provable in full MAV also has a normal proof. Therefore, the system with only analytic rules is complete for MAV provability. Horne \cite{Horne15:mav} proves this result via a syntactic proof involving rewriting and termination measures. In the following sections, we construct a semantic proof that normal proofs are complete for MAV.
